### 2021-09-06 data-accessibility.tex

\section*{Data Accessibility}

{\sc microhaplot} is available as an R package with these associated links:
\begin{itemize}
\item Stable version available on CRAN (when available): \url{https://CRAN.R-project.org/package=gscramble}
\item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
\item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
\item Data repository with full data from pig, Chinook, and bird (maybe?) data sets \url{https://dryad.something.or.other}
\end{itemize}
\mbox{}