### 2023-10-20 author-title-etc.tex
add matt to authors.  tiny edits

%!TEX root = main.tex

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{{\sc gscramble:} Simulation of admixed individuals without reuse of genetic material}
\title{\myTitle}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachel M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
\author{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachel M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{$^*$Fisheries Ecology Division,
    Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,
    110 McAllister Road,
    Santa Cruz, CA 95060, USA.,
    $^\dagger$Dept. of Fish, Wildlife, and Conservation Biology, Colorado State University, Fort Collins, CO, USA,
    $^\ddag$Dept. of Biology, Colorado State University, Fort Collins, CO, USA,
    $^\#$National Wildlife Research Center, United States Department of Agriculture, Wildlife Services,
    Fort Collins, CO, USA
}
\renewcommand{\AuthorAddresses}{\myAffiliations}

\renewcommand{\KeyWords}{genetic stock identification, mixture deconvolution, permutation testing, power analysis, blah, blah}

\renewcommand{\CorrespondingAuthor}{Eric C. Anderson, eric.anderson@noaa.gov}

% the email address for the corresponding author
\newcommand{\myEmailAddress}{eric.anderson@noaa.gov}
\newcommand{\myEmailFootnote}{$^\S$}

% here you can put your very own copyright notice
\newcommand{\myCopyright}{\copyright US Federal Government work in the public domain in the USA}

% here you can put a running title (a short title that goes on the left of the
% even pages)
\newcommand{\myRunningTitle}{Linkage-aware power assessment for admixture proportions}
\renewcommand{\RunningTitle}{\myRunningTitle}

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson {\em et al.}}

%%%% DONE WITH AUTHOR/TITLE/ETC INFORMATION DEFINITION