### 2024-02-05 abstract.tex
added draft abstract

%!TEX main.tex

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals 
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement, 
2) segregates alleles based on species-specific recombination rates. \gscramble{} is 
a flexible simulation approach that allows users to create pedigrees of varying complexity 
in order to simulate admixed genotypes and it allows users to track haplotype blocks 
from the source populations through the pedigrees. We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.