### 2024-07-11 data-accessibility.tex
Fixing quotation marks and urls...

\section*{Data Accessibility}

\begin{itemize}
\item Stable version of \gscramble{} available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}
\item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
\item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
\item Data and code repository with simulation analyses and steelhead-cutthroat analyses: \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} (To be transferred to Dryad or Zenodo upon acceptance).
\item Data and code repository for pig analyses: \url{https://github.com/mgdesaix/gscramble-paper-pigs} (To be transferred to Dryad or Zenodo upon acceptance)
\end{itemize}
\mbox{}