### 2024-07-12 data-accessibility.tex
Add author contributions

\section*{Author Contributions}

The gscramble simulation approach grew from discussions between TJS and ECA\@.
ECA, RMG, MGD, and TJS wrote the R package, `gscramble.'   Simulations of bias due to RISPI and
of variance in admixture fractions due to physical linkage were conducted by ECA.
The empirical analyses using `gscramble' were conducted
by MGD and RMG with data from TJS and by ECA with previously published data.  All authors contributed to writing and
editing the manuscript.

\section*{Data Availability Statement}

\begin{itemize}
\item Stable version of \gscramble{} available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}
\item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
\item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
\item Data and code repository with simulation analyses and steelhead-cutthroat analyses: \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} (To be transferred to Dryad or Zenodo upon acceptance).
\item Data and code repository for pig analyses: \url{https://github.com/mgdesaix/gscramble-paper-pigs} (To be transferred to Dryad or Zenodo upon acceptance)
\end{itemize}
\mbox{}