### 2024-12-03 data-accessibility.tex

\section*{Data Accessibility and Benefit-Sharing}

\subsection*{Data Accessibility Statement}

A stable version of the `gscramble' R package is available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}. The development version and entire revision history of
\gscramble{} is available on GitHub: \url{https://github.com/eriqande/gscramble}, and all package documentation is available online in HTML format: \url{https://eriqande.github.io/gscramble}

Data and code for the simulation analyses and steelhead-cutthroat analyses are available from \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} and are archived at \url{https://doi.org/10.5281/zenodo.14270660}.  Data and code for the pig analyses are available from \url{https://github.com/mgdesaix/gscramble-paper-pigs} and are archived
at \url{Get-from-Matt}.

\subsection*{Benefit-sharing Statement}

Benefits Generated: Benefits from this research accrue from the sharing of our data and results on public databases and the distribution of a novel and useful R package as described above.

\section*{Author Contributions}

The gscramble simulation approach grew from discussions between TJS and ECA\@.
ECA, RMG, MGD, and TJS wrote the R package, `gscramble.'   Simulations of bias due to RISPI and
of variance in admixture fractions due to physical linkage were conducted by ECA.
The empirical analyses using `gscramble' were conducted by MGD, RMG, TJS, and ECA with previously published data.
All authors contributed to writing and editing the manuscript.