### 2024-12-10 abstract.tex
Revised and resubmitted version

%!TEX root = main.tex

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals
(or the alleles that they carry)
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement and 
2) segregates alleles based on species-specific recombination rates. This framework
makes it possible to simulate admixed individuals in a way that respects the physical
linkage between markers on the same chromosome and which does not suffer
from RISPI.  This is achieved in \gscramble{} by allowing 
users to specify pedigrees of varying complexity 
in order to simulate admixed genotypes, segregating and tracking haplotype blocks 
from different source populations through those pedigrees, and then sampling---using
a variety of permutation schemes---alleles from empirical data into those haplotype blocks.
We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.